---
title: "Reproducible Research - Project 1"

output: html_document
---
By Ervachon - April 2015

###1 - Loading and preprocessing the data

Load the data (i.e. read.csv())  

```{r,echo = FALSE}
options(warn=-1)
ThePath <- "D:/_MOOC_/git/Reproducible_Research/"
```
NB: the variable **ThePath** is the complete path where is the activity file.
```{r}
ActivityMonotoringData <- read.csv(paste(ThePath ,"activity.csv",sep =""))
```

Process/transform the data (if necessary) into a format suitable for your analysis  
Here i omit the NA value.

```{r}
ActivityMonotoringDataWithoutNA <- na.omit(ActivityMonotoringData)
```

###2 - What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day.  
NB: don't forget to load the "plyr" library.

```{r}
library(plyr)
NumberOfStepsByDay <- ddply(ActivityMonotoringDataWithoutNA, 
                            .(ActivityMonotoringDataWithoutNA$date), 
                            summarise, sum = sum(steps, na.rm = TRUE))
hist(NumberOfStepsByDay$sum, 
     col = "blue", 
     main ="Number of steps by days",
     xlab ="Interval of the day",  
     ylab ="Number of steps")
```

Calculate and report the mean and median total number of steps taken per day
```{r}
mean(NumberOfStepsByDay$sum)
```
```{r}
median(NumberOfStepsByDay$sum)
```

###3 - What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
AverageStepsByInt <- ddply(ActivityMonotoringDataWithoutNA, .(interval), 
                            summarise, mean = mean(steps, na.rm = TRUE))
par(bg='white', fg='white', col='black', col.axis='magenta',
    col.lab='black', col.main='blue', col.sub='black') 
plot ( AverageStepsByInt$interval, 
       AverageStepsByInt$mean, 
       type = "l", 
       col = "red", 
       bg.col = "blue",
       main = "Daily activity",
       xlab ="Interval of the day",  
       ylab ="Number of steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
AverageStepsByInt[ AverageStepsByInt$mean == max(AverageStepsByInt$mean), ]
```

###4 - Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(ActivityMonotoringData$steps))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated, create a new dataset that is equal to the original dataset but with the missing data filled in.
   
   Here I choose to replace NA values with the mean and make a new dataset.
```{r}
ActivityMonotoringData2 <- ActivityMonotoringData
ActivityMonotoringData2$steps[is.na(ActivityMonotoringData$steps)] <- 
   AverageStepsByInt[AverageStepsByInt$interval == 
                        ActivityMonotoringData2$interval , ]$mean
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
NumberOfStepsByDay2 <- ddply(ActivityMonotoringData2, 
                             .(ActivityMonotoringData2$date), 
                             summarise, sum = sum(steps, na.rm = TRUE))

hist(NumberOfStepsByDay2$sum,
     col = "blue", 
     main ="Number of steps by days",
     xlab ="Interval of the day",  
     ylab ="Number of steps")
```
```{r}
mean(NumberOfStepsByDay2$sum)
```
```{r}
median(NumberOfStepsByDay2$sum)
```

Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

   The difference is that before the mean and the median were mostly identical (10766)  
   And now the difference was 9530 // 10439  
   we can notice that the median before and after is quit identical : 10766 // 10439  
   Instead of the mean : 10766 // 9530 so the median is more 
   
   

###5 - Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

i preferte use as.POSIXlt (and as.Date) Instead of weekdays because it return a integer 0 to 6 (sunday to saturday), easier to manipulate.

I create the new column
```{r}
ActivityMonotoringData2$typeDay <- "NA"
```
And i update the value :  
- 0,6 : it's sunday and saturday -> weekend  
- 1 to 5 : monday to friday > weekday  

```{r}
ActivityMonotoringData2$typeDay[
   which(as.POSIXlt(as.Date(ActivityMonotoringData2$date))$wday %in% c(0,6)) 
   ] <- "weekend"
ActivityMonotoringData2$typeDay[
   which(as.POSIXlt(as.Date(ActivityMonotoringData2$date))$wday %in% c(1,2,3,4,5)) 
   ] <- "weekday"
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
library(lattice)
AverageStepsByInt2 <- ddply(ActivityMonotoringData2, .(interval,typeDay), 
                           summarise, mean = mean(steps, na.rm = TRUE))

xyplot(  type = "l",
         data = AverageStepsByInt2,
         mean ~ interval | typeDay,
         main = "Mean of number of step by type of day",
         xlab = "Interval",
         ylab = "Mean of number of steps",
         layout = c(1,2),
         col.line = c("red")
)
```
  
### End of the document